Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-01-06 paper.tex
add relate curves

Similarly, we generated estimates of effective population size as inferred from
coalescence rates with Relate. Changes in Ne over time from prior studies
strongly support a bottleneck in Out of Africa populations, but African
populations also dip in effective size around 200kya, although this decrease is
more modest than the OOA event [Mallick 2016]. One possible explanation is that
ancestral population structure during the Middle Stone Age inflated Ne and
subsequent gene flow among ancestral Africans resulted in a lower Ne after
100-200kya [Chikhi?]. To our knowledge, explicit comparisons of this hypothesis
with SMC-style approaches have not been published. We simulated genomic data
under 4 models (Figure SX). All models, including the Single Origin model,
generate multi-modal curves indicating that the ancestral increase in Ne
between 100kya-1Mya is a general feature of the coalescent rate in humans.
Simulated Relate Ne curves under the CM and MM are distinguished by having 2
and 3 modes, respectively. The Continuous Migration model had the better
qualitative fit with the data, but the ancestral mode was shifted further back
in time relative to the data. All simulated models underestimated the level of
population growth during the Holocene, indicating additional free parameters
are necessary to refine these growth rates. \aprcomment{Discussion on this
paragraph moved to issues (issue \#1)}

>>>>

Similarly, we generated estimates of effective population size as inferred from
coalescence rates with Relate. Changes in Ne over time from prior studies
strongly support a bottleneck in Out of Africa populations, but African
populations also dip in effective size around 200kya, although this decrease is
more modest than the OOA event \citep{Mallick2016-lx}. One possible explanation
is that ancestral population structure during the Middle Stone Age inflated Ne
and subsequent gene flow among ancestral Africans resulted in a lower Ne after
100-200kya \citep{Mazet2016-wn}\aprcomment{Is this a good ref here?}. To our
knowledge, explicit comparisons of this hypothesis with SMC-style approaches
have not been published. We simulated genomic data under 4 models
\aprcomment{(Figure SX)}. All models, including the Single Origin model,
generate multi-modal curves indicating that the ancestral increase in Ne
between 100kya-1Mya is a general feature of the coalescent rate in humans.
Simulated Relate Ne curves under the CM and MM are distinguished by having 2
and 3 modes, respectively. The Continuous Migration model had the better
qualitative fit with the data, but the ancestral mode was shifted further back
in time relative to the data. All simulated models underestimated the level of
population growth during the Holocene, indicating additional free parameters
are necessary to refine these growth rates. \aprcomment{Discussion on this
paragraph moved to issues (issue \#1)}

### 2022-01-06 supplement.tex
add relate curves

>>>>

\begin{figure}[ht]
    \centering
    \includegraphics{figures/supp-relate-iicr-data}
    \caption{
        \textbf{Inverse instantaneous coalescence rates inferred from the data.}
    }
    \label{fig:supp-relate-sizes}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/supp-relate-iicr-simulation}
    \caption{
        \textbf{Inverse instantaneous coalescence rates reinferred from simulated data.}
    }
    \label{fig:supp-relate-sizes-sim}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/supp-relate-rccr-data}
    \caption{
        \textbf{Relative cross-coalescence rates inferred from the data.}
    }
    \label{fig:supp-relate-rccr}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/supp-relate-rccr-single-origin}
    \caption{
        \textbf{Relative cross-coalescence rates reinferred from simulated data
        under the single-origin model.}
    }
    \label{fig:supp-relate-rccr-sim}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/supp-relate-rccr-continuous-migration}
    \caption{
        \textbf{Relative cross-coalescence rates reinferred from simulated data
        under the continuous-migration model.}
    }
    \label{fig:supp-relate-rccr-sim}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/supp-relate-rccr-merger-without-stem-migration}
    \caption{
        \textbf{Relative cross-coalescence rates reinferred from simulated data
        under the merger-without-stem-migration-model.}
    }
    \label{fig:supp-relate-rccr-sim}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/supp-relate-rccr-merger-with-stem-migration}
    \caption{
        \textbf{Relative cross-coalescence rates reinferred from simulated data
        under the merger-with-stem-migration model.}
    }
    \label{fig:supp-relate-rccr-sim}
\end{figure}

\begin{figure}[ht]
    \centering
    %\includegraphics{figures/supp-deep-brances-data}
    \caption{
        \textbf{Neanderthal-matching deep branch distribution from the data}
    }
    \label{fig:supp-deep-branches}
\end{figure}

\begin{figure}[ht]
    \centering
    %\includegraphics{figures/supp-deep-branches-sim}
    \caption{
        \textbf{Deep branch distribution reinferred from simulated data.}
    }
    \label{fig:supp-deep-branches-sim}
\end{figure}

\begin{figure}[ht]
    \centering
    %\includegraphics{figures/supp-IM-misspecification}
    \caption{
        \textbf{IM model split times reinferred from simulated data of full model.}
    }
    \label{fig:supp-IM-misspecification}
\end{figure}
